---
title: "Profit Maximisation and Competitive Markets"
date: "September 29, 2015"
output: html_document
---

# 1. Taxes and Subsidies

The demand curve is given by

$$Q_{D} = 18 - 2P$$ 

The supply curve is given by

$$Q_{S} = -2 + 8P$$ 

```{r message=FALSE,warning=FALSE}

#load some libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)

#define the supply and demand functions
supply <- function(p) {
  return( -2 + 8*p )
}

demand <- function(p) {
  return(18-2*p)
}

supply_sub <- function(p,t) {
  return( -2 + 8*(p+t))
}

demand_sub <- function(p,t) {
  return(18-2*(p-t))
}

#create a dataframe with price going from 0 to 10, and calculate the other variables for each q
market1 <- data.frame(
  price=c(0:30)*0.1
  ) %>%
  mutate(
    supply = supply(price),
    demand = demand(price),
    supply_2 = supply_sub(price,1),
    demand_2 = demand_sub(price,1)
  )

market1Narrow <- market1 %>%
gather(variable,quantity,supply:demand_2)

```

Here's what the market looks like in equilibrium

```{r message=FALSE,warning=FALSE}

#plot value against quantity, grouping by variable
ggplot(filter(
  market1Narrow,variable!="demand_2" &variable!="supply_2"
  ),aes(quantity,price,colour=variable)) + geom_line() +xlim(c(0,30))

```

Here's what the data looks like
```{r message=FALSE,warning=FALSE}
kable(market1)
```

Here's a graphical representation of what happens if we give the subsidy to the supplier
```{r message=FALSE,warning=FALSE}

#plot value against quantity, grouping by variable
ggplot(filter(
  market1Narrow,variable!="demand_2"
  ),aes(quantity,price,colour=variable)) + geom_line() +xlim(c(0,30))

```


Here's a graphical representation of what happens if we give the subsidy to the consumer
```{r message=FALSE,warning=FALSE}

ggplot(filter(
  market1Narrow,variable!="supply_2"
  ),aes(quantity,price,colour=variable)) + geom_line() +xlim(c(0,30))

```

# 2. Profit Maximisation

Suppose that a firm's marginal cost is given by

$$MC(Q) = 3 + 2Q$$

and that price = 0

```{r}

#define the marginal cost function
marginal_cost <- function(q) {
  return(3 + 2*q )
}

#create a market table for the firm
market2 <- data.frame(
  quantity=c(0:10)
  ) %>%
  mutate(
    price = 9,
    marginal_cost = marginal_cost(quantity)
  )

market2Narrow <- market2 %>%
gather(variable,value,price:marginal_cost)

```

Here's what the data looks like
```{r message=FALSE,warning=FALSE}
kable(market2)
```

Here's a graphical representation of marginal costs agains price
```{r message=FALSE,warning=FALSE}

ggplot(market2Narrow,aes(quantity,value,colour=variable)) + geom_line() 

```


# 3. Market Power

Suppose that price is given by 

$$P=90-2Q$$  

Total cost is given by

$$Cost=100+2Q^2$$

Total revenue is given by

$$Revenue=P*Q=90Q-2Q^2$$

Marginal cost and marginal revenue are calculated by

$$Value_{q+1}-Value_q,$$

that is, the difference between the value at the current quantity and the value for an additional unit of quantity.

Profit is given by

$$Revenue-Cost$$

```{r message=FALSE,warning=FALSE}

#define the price and cost functions
price <- function(q) {
  return( 90 - 2*q )
}

cost <- function(q) {
  return(100+2*q^2)
}

#create a dataframe with quantity going from 0 to 20, and calculate the other variables for each q
market3 <- data.frame(
  quantity=c(0:20)
  ) %>%
  mutate(
    price = price(quantity),
    cost = cost(quantity),
    mar_cost = lead(cost)-cost,
    total_revenue_monopoly = price*quantity,
    marginal_revenue = lead(total_revenue_monopoly)-total_revenue_monopoly,
    profit_monopoly = total_revenue_monopoly - cost,
    average_costs = cost / quantity
    #simple_marginal_revenue = total_revenue-lag(total_revenue)
  )

```

In a table showing quantity from 1 to 20 we can see values for all our variables:

```{r}

kable(market3)

```


Here is what we happen when we plot them

```{r message=FALSE,warning=FALSE}
#gather the variables into two columns, variable and value
market3Narrow <- market3 %>%
gather(variable,value,price:average_costs)

#plot value against quantity, grouping by variable
ggplot(market3Narrow,aes(quantity,value,colour=variable)) + geom_line() 

#remove total revenue and cost for better scale
ggplot(filter(
  market3Narrow,variable != "total_revenue_monopoly" & variable != "cost"
  ), aes(quantity,value,colour=variable)) + geom_line()

ggplot(filter(
  market3Narrow,variable != "total_revenue_monopoly" & variable != "cost" & variable != "profit_monopoly"
  ), aes(quantity,value,colour=variable)) + geom_line()



```
